\vspace{1.0cm} \Large \bf Supplemental material of ``Health and disease imprinted in the time variability of the human microbiome''

### 2016-04-15 eLife/elife_sub.tex
Add eLife docs and ISMEJ letter to editor
On branch master
	new file:   ISMEJ/lettertoeditor_sub.docx
	deleted:    eLife/To_be_populated.txt
	new file:   eLife/elife_sub.tex
	new file:   eLife/elife_suppl_sub.tex

% Use only LaTeX2e, calling the article.cls class and 12-point type.

\documentclass[12pt]{article}

% Users of the {thebibliography} environment or BibTeX should use the
% scicite.sty package, downloadable from *Science* at
% www.sciencemag.org/about/authors/prep/TeX_help/ .
% This package should properly format in-text
% reference calls and reference-list numbers.

\usepackage{scicite}
\usepackage{graphicx}

% Use times if you have the font installed; otherwise, comment out the
% following line.

\usepackage{times}

% The preamble here sets up a lot of new/revised commands and
% environments.  It's annoying, but please do *not* try to strip these
% out into a separate .sty file (which could lead to the loss of some
% information when we convert the file to other formats).  Instead, keep
% them in the preamble of your main LaTeX source file.

% The following parameters seem to provide a reasonable page setup.

\topmargin 0.0cm
\oddsidemargin 0.2cm
\textwidth 16cm 
\textheight 21cm
\footskip 1.0cm

%The next command sets up an environment for the abstract to your paper.

\newenvironment{sciabstract}{%
\begin{quote} \bf}
{\end{quote}}

% If your reference list includes text notes as well as references,
% include the following line; otherwise, comment it out.

\renewcommand\refname{References and Notes}

% The following lines set up an environment for the last note in the
% reference list, which commonly includes acknowledgments of funding,
% help, etc.  It's intended for users of BibTeX or the {thebibliography}
% environment.  Users who are hand-coding their references at the end
% using a list environment such as {enumerate} can simply add another
% item at the end, and it will be numbered automatically.

\newcounter{lastnote}
\newenvironment{scilastnote}{%
\setcounter{lastnote}{\value{enumiv}}%
\addtocounter{lastnote}{+1}%
\begin{list}%
{\arabic{lastnote}.}
{\setlength{\leftmargin}{.22in}}
{\setlength{\labelsep}{.5em}}}
{\end{list}}

% Include your paper's title here

\title{Microbiota: are you sick?}

% Place the author information here.  Please hand-code the contact
% information and notecalls; do *not* use \footnote commands.  Let the
% author contact information appear immediately below the author names
% as shown.  We would also prefer that you don't change the type-size
% settings shown here.

\author
{Jose Manuel Mart\'i,$^{1}$ Daniel M. Mart\'inez,$^{2}$ Manuel Pe\~na$^{1}$, C\'esar Gracia$^{1}$, \\
Amparo Latorre$^{2,3,4}$, Andr\'es Moya$^{2,3,4}$ \& Carlos P. Garay$^{1,\ast}$\\
\\
\normalsize{$^{1}$Instituto de Fisica Corpuscular, CSIC-UVEG, P.O.  22085, 46071, Valencia, Spain.}\\
\normalsize{$^{2}$FISABIO, Avda de Catalunya, 21, 46020, Valencia, Spain.}\\
\normalsize{$^{3}$Cavanilles Institute of Biodiversity and Evolutionary Biology, Univ. de Valencia, 46980, Spain.}\\
\normalsize{$^{4}$CIBER en Epidemiolog\' ia y Salud P\'ublica (CIBEResp), Madrid, Spain.}\\
\\
\normalsize{$^\ast$To whom correspondence should be addressed; E-mail:  penya@ific.uv.es}
}

% Include the date command, but leave its argument blank.

@@

%%%%%%%%%%%%%%%%% END OF PREAMBLE %%%%%%%%%%%%%%%%

@@

% Double-space the manuscript.

\baselineskip24pt

% Make the title.

@@

% Place your abstract within the special {sciabstract} environment.

\begin{sciabstract}
Human microbiota plays an important role in determining changes from health to disease. Increasing research activity is dedicated to understand its diversity and variability. 
We analyse 16S rRNA and whole genome sequencing (WGS) data from the gut microbiota of 97 individuals monitored in time. Temporal fluctuations in the microbiome reveal 
significant differences due to factors that affect the microbiota such as dietary changes, antibiotic intake, early gut development or disease. Here we show that a fluctuation 
scaling law describes the temporal variability of the system and that a noise-induced phase transition is central in the route to disease. The universal law distinguishes healthy 
from sick microbiota and quantitatively characterizes the path in the phase space, which opens up its potential clinical use and, more generally, other technological 
applications where microbiota plays an important role. 
\end{sciabstract}

% In setting up this template for *Science* papers, we've used both
% the \section* command and the \paragraph* command for topical
% divisions.  Which you use will of course depend on the type of paper
% you're writing.  Review Articles tend to have displayed headings, for
% which \section* is more appropriate; Research Articles, when they have
% formal topical divisions at all, tend to signal them with bold text
% that runs into the paragraph, for which \paragraph* is the right
% choice.  Either way, use the asterisk (*) modifier, as shown, to
% suppress numbering.

\section*{Introduction}

@@

@@

@@

Taylor's power law has been explained in terms of various effects, all without general consensus. It can be shown to have its origin in a mathematical convergence similar to the central limit theorem, so virtually any statistical model designed to produce a Taylor law converge to a Tweedie distribution\cite{stat}, providing a mechanistic explanation based on the statistical theory of errors\cite{convergence1,convergence2,convergence3}. To unveil the generic mechanisms that drive different scenarios in the $\beta$--V space, we model the system by assuming that taxon relative abundance follows a Langevin equation with a deterministic term that captures the fitness of each taxon and a randomness term with Gaussian random noise\cite{ranking}. Both terms are modelled by power laws, with coefficients that can be interpreted as the taxon fitness $F_i$ and the variability $V$ (see Section 1 in supplemental material). When $V$ is sufficiently low, abundances are stable in time.  Differences in variability $V$ can induce a noise-induced phase transition in relative abundances of taxa. The temporal evolution of the probability of a taxon having abundance $x_i$ given its fitness is governed by the Fokker--Planck equation. The results of solving this equation show that  the stability is best captured by fitness F and amplitude of fluctuations V phase space (see Figure 2). 
The model predicts two phases for the gut microbiome: a stable phase with large variability that permits some changes in the relative abundances of taxa and an unstable phase with larger variability, above the phase transition, where the order of abundant taxa varies significantly with time. The microbiome of all healthy individuals was found to be in the stable phase, while the microbiome of several other individuals was shown to be in the unstable phase. In particular, individuals taking antibiotics and IBS--diagnosed patient P2 had the most severe symptoms. In this phase diagram, each microbiota state is represented by a point at its measured variability V and inferred fitness F. The model predicts high average fitness for all taxa, i.e., taxa are narrowly distributed in F. The fitness parameter has been chosen with different values for demonstrative purposes. Fitness is larger for the healthiest subjects and smaller for the IBS--diagnosed patients.

In summary, we have quantitatively characterized whether the microbiota belongs to a healthy individual or a subject corresponding to an altered or pathological state (ie, altered diet, antibiotic treatment, early gut development, diagnosed IBS). Deciphering the mechanisms of disease requires in depth knowledge of the underlying biological mechanisms. We describe here the macroscopic behavior of disease by a noise-induced phase transition with a control parameter that can be measured by the temporal variability of the microbiome. The microbiota of healthy individuals and of individuals with pathologies represent different phases separated by this noise-induced phase transition. Improved high--throughput sequencing of samples from individuals monitored over time and taxonomic assigning methods will provide a better distinction among pathologies or altered states of the microbiota.

>>>>

\documentclass[11pt,oneside,letterpaper]{article}

% graphicx package, useful for including eps and pdf graphics
\usepackage{graphicx}
\DeclareGraphicsExtensions{.pdf,.png,.jpg}

% basic packages
\usepackage{color} 
\usepackage{parskip}
\usepackage{float}

% text layout
\usepackage{geometry}
\geometry{textwidth=15cm} % 15.25cm for single-space, 16.25cm for double-space
\geometry{textheight=22cm} % 22cm for single-space, 22.5cm for double-space

% helps to keep figures from being orphaned on a page by themselves
\renewcommand{\topfraction}{0.85}
\renewcommand{\textfraction}{0.1}

% bold the 'Figure #' in the caption and separate it with a period
% Captions will be left justified
\usepackage[labelfont=bf,labelsep=period,font=small]{caption}

% review layout with double-spacing
\usepackage{setspace} 
\doublespacing
\captionsetup{labelfont=bf,labelsep=period,font=doublespacing}

% cite package, to clean up citations in the main text. Do not remove.
\usepackage{cite}
\renewcommand\citeleft{(}
\renewcommand\citeright{)}
\renewcommand\citeform[1]{\textsl{#1}}

% Remove brackets from numbering in list of References
\renewcommand\refname{\large References}
\makeatletter
\renewcommand{\@biblabel}[1]{\quad#1.}
\makeatother

\usepackage{authblk}
\renewcommand\Authands{ \& }
\renewcommand\Authfont{\normalsize \bf}
\renewcommand\Affilfont{\small \normalfont}
\makeatletter
\renewcommand\AB@affilsepx{, \protect\Affilfont}
\makeatother

% notation
\usepackage{amsmath}
\usepackage{amssymb}

%%% TITLE %%%
\title{\vspace{1.0cm} \Large \bf 
Health and disease imprinted in the time variability of the human microbiome
}

\author[1]{Jose Manuel Mart\' i}
\author[1,2]{Daniel M. Mart\' inez}
\author[1]{Manuel Pe\~na}
\author[1]{C\'esar Gracia}
\author[2,3,4]{Amparo Latorre}
\author[2,3,4]{Andr\'es Moya}
\author[1]{Carlos P. Garay}

\affil[1]{Instituto de Fisica Corpuscular, CSIC-UVEG, P.O.  22085, 46071, Valencia, Spain.}
\affil[2]{FISABIO, Avda de Catalunya, 21, 46020, Valencia, Spain.}
\affil[3]{Cavanilles Institute of Biodiversity and Evolutionary Biology, Univ. de Valencia, 46980, Spain. }
\affil[4]{CIBER en Epidemiolog\' ia y Salud P\'ublica (CIBEResp), Madrid, Spain}

@@

@@

@@

%%% ABSTRACT %%%
\begin{abstract}

Human microbiota plays an important role in determining changes from health to disease. Increasing research activity is dedicated to understand its diversity and variability. 
We analyse 16S rRNA and whole genome sequencing (WGS) data from the gut microbiota of 97 individuals monitored in time. Temporal fluctuations in the microbiome reveal 
significant differences due to factors that affect the microbiota such as dietary changes, antibiotic intake, early gut development or disease. Here we show that a fluctuation 
scaling law describes the temporal variability of the system and that a noise-induced phase transition is central in the route to disease. The universal law distinguishes healthy 
from sick microbiota and quantitatively characterizes the path in the phase space, which opens up its potential clinical use and, more generally, other technological 
applications where microbiota plays an important role.

\end{abstract}

\pagebreak

%%% INTRODUCTION %%%
\section*{Introduction}

@@

%%% RESULTS %%%
\section*{Results and discussion}

@@

Taylor's power law has been explained in terms of various effects, all without general consensus. It can be shown to have its origin in a mathematical convergence similar to the central limit theorem, so virtually any statistical model designed to produce a Taylor law converge to a Tweedie distribution\cite{stat}, providing a mechanistic explanation based on the statistical theory of errors\cite{convergence1,convergence2,convergence3}. To unveil the generic mechanisms that drive different scenarios in the $\beta$--V space, we model the system by assuming that taxon relative abundance follows a Langevin equation with a deterministic term that captures the fitness of each taxon and a randomness term with Gaussian random noise\cite{ranking}. Both terms are modelled by power laws, with coefficients that can be interpreted as the taxon fitness $F_i$ and the variability $V$ (see Section 1 in supplemental material). When $V$ is sufficiently low, abundances are stable in time.  Differences in variability $V$ can induce a noise-induced phase transition in relative abundances of taxa. The temporal evolution of the probability of a taxon having abundance $x_i$ given its fitness is governed by the Fokker--Planck equation. The results of solving this equation show that  the stability is best captured by fitness F and amplitude of fluctuations V phase space (see Figure 2).

@@

The model predicts two phases for the gut microbiome: a stable phase with large variability that permits some changes in the relative abundances of taxa and an unstable phase with larger variability, above the phase transition, where the order of abundant taxa varies significantly with time. The microbiome of all healthy individuals was found to be in the stable phase, while the microbiome of several other individuals was shown to be in the unstable phase. In particular, individuals taking antibiotics and IBS--diagnosed patient P2 had the most severe symptoms. In this phase diagram, each microbiota state is represented by a point at its measured variability V and inferred fitness F. The model predicts high average fitness for all taxa, i.e., taxa are narrowly distributed in F. The fitness parameter has been chosen with different values for demonstrative purposes. Fitness is larger for the healthiest subjects and smaller for the IBS--diagnosed patients.

\section*{Conclusions}

We have quantitatively characterized whether the microbiota belongs to a healthy individual or a subject corresponding to an altered or pathological state (ie, altered diet, antibiotic treatment, early gut development, diagnosed IBS). Deciphering the mechanisms of disease requires in depth knowledge of the underlying biological mechanisms. We describe here the macroscopic behavior of disease by a noise-induced phase transition with a control parameter that can be measured by the temporal variability of the microbiome. The microbiota of healthy individuals and of individuals with pathologies represent different phases separated by this noise-induced phase transition. Improved high--throughput sequencing of samples from individuals monitored over time and taxonomic assigning methods will provide a better distinction among pathologies or altered states of the microbiota.

### 2016-04-15 eLife/elife_suppl_sub.tex
Add eLife docs and ISMEJ letter to editor
On branch master
	new file:   ISMEJ/lettertoeditor_sub.docx
	deleted:    eLife/To_be_populated.txt
	new file:   eLife/elife_sub.tex
	new file:   eLife/elife_suppl_sub.tex

% Use only LaTeX2e, calling the article.cls class and 12-point type.

\documentclass[12pt]{article}

% Users of the {thebibliography} environment or BibTeX should use the
% scicite.sty package, downloadable from *Science* at
% www.sciencemag.org/about/authors/prep/TeX_help/ .
% This package should properly format in-text
% reference calls and reference-list numbers.

\usepackage{scicite}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage[usenames,dvipsnames,table]{xcolor}
\usepackage[nolists,tablesfirst, nomarkers]{endfloat}

% Use times if you have the font installed; otherwise, comment out the
% following line.

\usepackage{times}

% The preamble here sets up a lot of new/revised commands and
% environments.  It's annoying, but please do *not* try to strip these
% out into a separate .sty file (which could lead to the loss of some
% information when we convert the file to other formats).  Instead, keep
% them in the preamble of your main LaTeX source file.

% The following parameters seem to provide a reasonable page setup.

\topmargin 0.0cm
\oddsidemargin 0.2cm
\textwidth 16cm 
\textheight 21cm
\footskip 1.0cm

%The next command sets up an environment for the abstract to your paper.

\newenvironment{sciabstract}{%
\begin{quote} \bf}
{\end{quote}}

% If your reference list includes text notes as well as references,
% include the following line; otherwise, comment it out.

\newcommand{\CC}[0]{\emph{cmplxcruncher}}
\newcommand{\task}[1]{\texttt{\bfseries\scshape\textcolor{MidnightBlue}{#1}}}
\newcommand{\un}[1]{\operatorname{#1}}
\newcommand{\unclassrate}[0]{\un{kpb/s/core}}
\renewcommand\refname{References and Notes}
\renewcommand{\thefootnote}{\Roman{footnote}}

% The following lines set up an environment for the last note in the
% reference list, which commonly includes acknowledgments of funding,
% help, etc.  It's intended for users of BibTeX or the {thebibliography}
% environment.  Users who are hand-coding their references at the end
% using a list environment such as {enumerate} can simply add another
% item at the end, and it will be numbered automatically.

\newcounter{lastnote}
\newenvironment{scilastnote}{%
\setcounter{lastnote}{\value{enumiv}}%
\addtocounter{lastnote}{+1}%
\begin{list}%
{\arabic{lastnote}.}
{\setlength{\leftmargin}{.22in}}
{\setlength{\labelsep}{.5em}}}
{\end{list}}

% Include your paper's title here

\title{Supplemental material of ``Microbiota: are you sick?''}

% Place the author information here.  Please hand-code the contact
% information and notecalls; do *not* use \footnote commands.  Let the
% author contact information appear immediately below the author names
% as shown.  We would also prefer that you don't change the type-size
% settings shown here.

\author
{Jose Manuel Mart\'i,$^{1}$ Daniel M. Mart\'inez,$^{2}$ Manuel Pe\~na$^{1}$, C\'esar Gracia$^{1}$, \\
Amparo Latorre$^{2,3,4}$, Andr\'es Moya$^{2,3,4}$ \& Carlos P. Garay$^{1,\ast}$\\
\\
\normalsize{$^{1}$Instituto de Fisica Corpuscular, CSIC-UVEG, P.O.  22085, 46071, Valencia, Spain.}\\
\normalsize{$^{2}$FISABIO, Avda de Catalunya, 21, 46020, Valencia, Spain.}\\
\normalsize{$^{3}$Cavanilles Institute of Biodiversity and Evolutionary Biology, Univ. de Valencia, 46980, Spain.}\\
\normalsize{$^{4}$CIBER en Epidemiolog\' ia y Salud P\'ublica (CIBEResp), Madrid, Spain.}\\
\\
\normalsize{$^\ast$To whom correspondence should be addressed; E-mail:  penya@ific.uv.es}
}

% Include the date command, but leave its argument blank.

@@

% In setting up this template for *Science* papers, we've used both
% the \section* command and the \paragraph* command for topical
% divisions.  Which you use will of course depend on the type of paper
% you're writing.  Review Articles tend to have displayed headings, for
% which \section* is more appropriate; Research Articles, when they have
% formal topical divisions at all, tend to signal them with bold text
% that runs into the paragraph, for which \paragraph* is the right
% choice.  Either way, use the asterisk (*) modifier, as shown, to
% suppress numbering.

\section*{Model} \label{sec:model}

@@

        \item If $\left|\mathrm{AIC_{c_{NLR}}} - \mathrm{AIC_{c_{LLR}}}\right| \leq 2$, no model is favoured, so proceed with averaging:

@@

\newpage
\begin{thebibliography}{1}
	\bibitem{ranking} Blumm, N. \textit{et al.} Dynamics of ranking processes in complex systems. {\it Phys. Rev. Lett.} {\bf 109,} 128701 (2012).
	\bibitem{FD} Weber, J. \textit{et al.} Fluctuation dissipation theorem. {\it Phys. Rev.} {\bf 101}, 1620-6 (1956).
	\bibitem{kwashiorkor} Smith M.I.  \textit{et al.} Gut microbiomes of Malawian twin pairs discordant for kwashiorkor. {\it Science} {\bf 339,} 548-54 (2013).
	\bibitem{FASTX} Gordon, A., Hannon, G.J. FASTX-Toolkit. FASTQ/A shortreads pre-processing tools (2010). http://hannonlab.cshl.edu/fastx\_toolkit/ (accessed 23 Feb 2015).
	\bibitem{QIIME} Caporaso, J.G. \textit{et al.} QIIME allows analysis of high-throughput community sequencing data. {\it Nature Methods} {\bf 7,} 335-6 (2010).
	\bibitem{SILVA} Quast C.  \textit{et al.} The SILVA ribosomal RNA gene database project: improved data processing and web-based tools (2013)
	\bibitem{LMAT} Ames, S.K.  \textit{et al.} Scalable metagenomic taxonomy classification usng a reference genome database. {\it Bioinformatics} {\bf 29,} 2253-2260 (2013).
	\bibitem{antibiotic} Dethlefsen L., Relman D. A. Incomplete recovery and individualized responses of the human distal gut microbiota to repeated antibiotic perturbation. {\it Proc. Nat. Acad. Sci. USA} {\bf 108,} 4554-61 (2011).
	\bibitem{diet} David, L.A. \textit{et al.} Diet rapidly and reproducibly alters the human gut microbiome. {\it Nature} {\bf 505,} 559-63 (2014).
	\bibitem{IBS} Durban, A. \textit{et al.} Structural alterations of faecal and mucosa-associated bacterial communities in irritable bowel syndrome. {\it FEMS Microbiol. Ecol.} {\bf 86}, 581-9 (2013).
	\bibitem{LEA} Faith, J.J.  \textit{et al.} The long-term stability of the human gut microbiota. {\it Science} {\bf 341,} 1237439 (2013).
	\bibitem{moving} Caporaso, J.G.  \textit{et al.} Moving pictures of the human microbiome. {\it Genome Biol.} {\bf 12,} R50 (2011).
 	\bibitem{ecology} Xiao Xiao, Ethan P. White, Mevin B. Hooten, and Susan L. Durham. On the use of log-transformation vs. nonlinear regression for analyzing biological power laws. {\it Ecology} {\bf 92}, 10, 1887-1894 (2011).
 	\bibitem{genR2} Magee L., $R^2$ measures based on wald and likelihood ratio joint significance tests. {\it The American Statistician} {\bf 44}, 3, 250-253 (1990).
	\bibitem{disR2} Nagelkerke N.J.D., A note on a general definition of the coefficient of determination. {\it Biometrika} {\bf 78}, 3, 691-692 (1991).
	\bibitem{boot} Wu, C.F.J. Jackknife, bootstrap and other resampling methods in regression analysis. (with discussions) \textit{The Annals of Statistics} {\bf 14}: 1261?1350 (1986)
\end{thebibliography}

>>>>

\documentclass[11pt,oneside,letterpaper]{article}

% graphicx package, useful for including eps and pdf graphics
\usepackage{graphicx}
\DeclareGraphicsExtensions{.pdf,.png,.jpg}

% basic packages
\usepackage{color} 
\usepackage{parskip}
\usepackage{float}

% text layout
\usepackage{geometry}
\geometry{textwidth=15cm} % 15.25cm for single-space, 16.25cm for double-space
\geometry{textheight=22cm} % 22cm for single-space, 22.5cm for double-space

% helps to keep figures from being orphaned on a page by themselves
\renewcommand{\topfraction}{0.85}
\renewcommand{\textfraction}{0.1}
%% Roman footnote numbering style to not collide with bibliographic cites: 
\renewcommand{\thefootnote}{\Roman{footnote}}

\newcommand{\CC}[0]{\emph{cmplxcruncher}}
\newcommand{\task}[1]{\texttt{\bfseries\scshape\textcolor{MidnightBlue}{#1}}}
\newcommand{\un}[1]{\operatorname{#1}}
\newcommand{\unclassrate}[0]{\un{kpb/s/core}}

% bold the 'Figure #' in the caption and separate it with a period
% Captions will be left justified
\usepackage[labelfont=bf,labelsep=period,font=small]{caption}

% review layout with double-spacing
\usepackage{setspace} 
\doublespacing
\captionsetup{labelfont=bf,labelsep=period,font=doublespacing}

% cite package, to clean up citations in the main text. Do not remove.
\usepackage{cite}
\renewcommand\citeleft{(}
\renewcommand\citeright{)}
\renewcommand\citeform[1]{\textsl{#1}}

% Remove brackets from numbering in list of References
\renewcommand\refname{\large References}
\makeatletter
\renewcommand{\@biblabel}[1]{\quad#1.}
\makeatother

\usepackage{authblk}
\renewcommand\Authands{ \& }
\renewcommand\Authfont{\normalsize \bf}
\renewcommand\Affilfont{\small \normalfont}
\makeatletter
\renewcommand\AB@affilsepx{, \protect\Affilfont}
\makeatother

% notation
\usepackage{amsmath}
\usepackage{amssymb}

%%% TITLE %%%
\title{\vspace{1.0cm} \Large \bf 
Supplemental material of ``Health and disease imprinted in the time variability of the human microbiome''
}

\author[1]{Jose Manuel Mart\' i}
\author[1,2]{Daniel M. Mart\' inez}
\author[1]{Manuel Pe\~na}
\author[1]{C\'esar Gracia}
\author[2,3,4]{Amparo Latorre}
\author[2,3,4]{Andr\'es Moya}
\author[1]{Carlos P. Garay}

\affil[1]{Instituto de Fisica Corpuscular, CSIC-UVEG, P.O.  22085, 46071, Valencia, Spain.}
\affil[2]{FISABIO, Avda de Catalunya, 21, 46020, Valencia, Spain.}
\affil[3]{Cavanilles Institute of Biodiversity and Evolutionary Biology, Univ. de Valencia, 46980, Spain. }
\affil[4]{CIBER en Epidemiolog\' ia y Salud P\'ublica (CIBEResp), Madrid, Spain}

@@

\section{Model} \label{sec:model}

@@

        \item If $\left|\mathrm{AIC_{c_{NLR}}} - \mathrm{AIC_{c_{LLR}}}\right| \leq 2$, no model is favoured, so proceed with model averaging:

@@

\newpage
\begin{thebibliography}{99}
	\bibitem{ranking} Blumm, N. \textit{et al.} Dynamics of ranking processes in complex systems. {\it Phys. Rev. Lett.} {\bf 109,} 128701 (2012).
	\bibitem{FD} Weber, J. \textit{et al.} Fluctuation dissipation theorem. {\it Phys. Rev.} {\bf 101}, 1620-6 (1956).
	\bibitem{kwashiorkor} Smith M.I.  \textit{et al.} Gut microbiomes of Malawian twin pairs discordant for kwashiorkor. {\it Science} {\bf 339,} 548-54 (2013).
	\bibitem{FASTX} Gordon, A., Hannon, G.J. FASTX-Toolkit. FASTQ/A shortreads pre-processing tools (2010). http://hannonlab.cshl.edu/fastx\_toolkit/ (accessed 23 Feb 2015).
	\bibitem{QIIME} Caporaso, J.G. \textit{et al.} QIIME allows analysis of high-throughput community sequencing data. {\it Nature Methods} {\bf 7,} 335-6 (2010).
	\bibitem{SILVA} Quast C.  \textit{et al.} The SILVA ribosomal RNA gene database project: improved data processing and web-based tools (2013)
	\bibitem{LMAT} Ames, S.K.  \textit{et al.} Scalable metagenomic taxonomy classification usng a reference genome database. {\it Bioinformatics} {\bf 29,} 2253-2260 (2013).
	\bibitem{antibiotic} Dethlefsen L., Relman D. A. Incomplete recovery and individualized responses of the human distal gut microbiota to repeated antibiotic perturbation. {\it Proc. Nat. Acad. Sci. USA} {\bf 108,} 4554-61 (2011).
	\bibitem{diet} David, L.A. \textit{et al.} Diet rapidly and reproducibly alters the human gut microbiome. {\it Nature} {\bf 505,} 559-63 (2014).
	\bibitem{IBS} Durban, A. \textit{et al.} Structural alterations of faecal and mucosa-associated bacterial communities in irritable bowel syndrome. {\it FEMS Microbiol. Ecol.} {\bf 86}, 581-9 (2013).
	\bibitem{LEA} Faith, J.J.  \textit{et al.} The long-term stability of the human gut microbiota. {\it Science} {\bf 341,} 1237439 (2013).
	\bibitem{moving} Caporaso, J.G.  \textit{et al.} Moving pictures of the human microbiome. {\it Genome Biol.} {\bf 12,} R50 (2011).
 	\bibitem{ecology} Xiao Xiao, Ethan P. White, Mevin B. Hooten, and Susan L. Durham. On the use of log-transformation vs. nonlinear regression for analyzing biological power laws. {\it Ecology} {\bf 92}, 10, 1887-1894 (2011).
 	\bibitem{genR2} Magee L., $R^2$ measures based on wald and likelihood ratio joint significance tests. {\it The American Statistician} {\bf 44}, 3, 250-253 (1990).
	\bibitem{disR2} Nagelkerke N.J.D., A note on a general definition of the coefficient of determination. {\it Biometrika} {\bf 78}, 3, 691-692 (1991).
	\bibitem{boot} Wu, C.F.J. Jackknife, bootstrap and other resampling methods in regression analysis. (with discussions) \textit{The Annals of Statistics} {\bf 14}: 1261?1350 (1986)
\end{thebibliography}
